\documentclass[10pt]{article}
\usepackage[margin=0.9in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{caption}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\captionsetup{font=small}
\setlength{\parskip}{4pt}
\setlength{\parindent}{0pt}

\title{\vspace{-2.5em}Tuning DALS: An 80-Trial Hyperparameter Search\\
\large Can the level-set hyperparameters move the needle, or is the ceiling elsewhere?}
\author{CS269 Project --- Deformable Models}
\date{\today}

\begin{document}
\maketitle
\vspace{-1.5em}

\section{Question}

DALS couples a U-Net trunk to a differentiable Chan--Vese level-set evolution
with learnable parameters ($\mu$, $\lambda_1$, $\lambda_2$, $dt$), unrolled over
$n_{\text{iter}}$ steps. The structural and loss hyperparameters were never
searched --- they sit at hand-picked defaults. This mini report asks a simple
question: \emph{does any hyperparameter configuration beat the current DALS
baselines (FLAME-1 IoU 0.747, FLAME-3 IoU 0.122) beyond training noise?} A second
question rides along: does optimizing for \emph{boundary} fidelity select a
different configuration than optimizing for \emph{region} overlap?

\section{Method}

Random search, four arms = $\{$FLAME-3, FLAME-1$\} \times \{$val-IoU objective,
val-BF objective$\}$, 20 trials each (80 trials total). Each trial trains DALS at
reduced epochs (FLAME-3: 25; FLAME-1: 12 at 1024\,px), records \emph{both} val IoU
and val boundary F-score@2px, and the arm ranks by its own objective. The winner
of each arm is retrained at the full 50 epochs and scored on the held-out test
split (94 FLAME-3 / 301 FLAME-1 frames).

\paragraph{Search space.} $n_{\text{iter}}\!\in\!\{3,5,8,12\}$;
$\mu\!\in\![0.1,1.0]$; $\lambda_1,\lambda_2\!\in\![0.5,2.0]$;
$dt\!\in\![0.05,0.3]$; loss $\in\{$bce\_dice, focal\_tversky$\}$ with Focal--Tversky
$\beta\!\in\![0.5,0.8]$ ($\alpha\!=\!1\!-\!\beta$), $\gamma\!\in\![0.5,1.0]$.
Fixed at defaults: $\text{lr}=10^{-4}$, batch $=4$, light augmentation. Config
sampling is seeded for reproducibility.

\section{Results}

Every one of the four arms ranked the \emph{same} configuration first:
\[
n_{\text{iter}}\!=\!3,\;\mu\!=\!0.21,\;\lambda_1\!=\!0.93,\;\lambda_2\!=\!1.38,\;
dt\!=\!0.19,\;\text{loss}=\text{bce\_dice}.
\]

\begin{table}[H]
\centering
\caption{Winning config retrained at 50 epochs, test-split evaluation. Baseline
is the existing DALS checkpoint.}
\begin{tabular}{llrrr}
\toprule
Dataset & Search objective & Baseline IoU & Tuned test IoU & Tuned test Dice \\
\midrule
FLAME-3 & val IoU & 0.122 & 0.120 & 0.184 \\
FLAME-3 & val BF  & 0.122 & \textbf{0.129} & 0.193 \\
FLAME-1 & val IoU & 0.747 & \textbf{0.753} & 0.858 \\
FLAME-1 & val BF  & 0.747 & 0.745 & 0.853 \\
\bottomrule
\end{tabular}
\end{table}

\section{Findings}

\paragraph{No region-vs-boundary tradeoff.} On both datasets the val-IoU arm and
the val-BF arm selected the \emph{identical} winning hyperparameters. Optimizing
for boundary fidelity does not pull the DALS configuration anywhere different than
optimizing for region overlap --- there is no Pareto frontier to exploit.

\paragraph{FLAME-3 is a wash.} The tuned test IoU (0.120 / 0.129) straddles the
0.122 baseline. Tellingly, two retrains of the \emph{same} config differed by
0.009 IoU from training stochasticity alone (no fixed retrain seed), so the
run-to-run noise floor is $\approx\!\pm0.01$ --- the entire tuned-vs-baseline gap
sits inside it. This is consistent with the detection-bottleneck story: on
FLAME-3 the ground truth is hidden $\ge150^\circ$C thermal support, and the
ceiling is set by what the U-Net trunk can \emph{detect}, not by the level-set
hyperparameters.

\paragraph{FLAME-1 is a wash too, but confirms a strong baseline.} Best tuned test
IoU 0.753 vs.\ 0.747 baseline is a $+0.006$ nudge, again within noise. Tuning
neither broke nor meaningfully improved the already-strong visible-flame regime.

\paragraph{$n_{\text{iter}}\!=\!3$ and bce\_dice dominate.} Across all arms the
top trials clustered on few level-set iterations and the bce\_dice loss;
high-$n_{\text{iter}}$ and Focal--Tversky trials were frequently unstable (several
near-zero val IoU). The default DALS recipe is already close to optimal.

\section{Bottom line}

The 80-trial search found no hyperparameter configuration that beats the existing
DALS baselines beyond training noise on either dataset. The defaults are
effectively already tuned. If the goal is to actually move the FLAME-3 number, the
lever is the detection trunk --- recovering more of the hidden thermal support ---
not the contour hyperparameters.

\end{document}
