\documentclass[10pt]{article}
\usepackage[margin=0.85in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{caption}
\usepackage[hidelinks]{hyperref}
\usepackage{float}
\usepackage{xcolor}
\captionsetup{font=small}
\setlength{\parskip}{4pt}
\setlength{\parindent}{0pt}

\title{\vspace{-2.5em}Deep Snake (Paper-Faithful) in Context\\
\large Cross-method comparison on FLAME-3 and FLAME-1}
\author{CS269 Project --- Deformable Models}
\date{\today}

\begin{document}
\maketitle
\vspace{-1.5em}

\section{What this compares}

We place the new paper-faithful \texttt{deep\_snake\_paper} (separately-trained
CenterNet detector $\to$ octagon init $\to$ circular-convolution snake) against
every other method on both datasets. ``Deep Snake (simple)'' is the original
detector-free variant (U-Net coarse mask seeds the contour). All scores are mean
IoU/Dice on binary masks; rows are sorted by IoU.

\textbf{Two datasets, two definitions of ``fire'':}
\begin{itemize}
\item \textbf{FLAME-3} (Sycan Marsh): GT = paired thermal frame thresholded at
  $\ge150\,^\circ$C. A smoke-occluded hot region, largely \emph{invisible} as
  orange flame in RGB.
\item \textbf{FLAME-1}: GT = hand-labeled \emph{visible-flame} PNG masks. The label
  is the flame you can see in RGB.
\end{itemize}
All methods take RGB only as input.

\section{FLAME-3 (thermal-defined fire)}

\begin{table}[H]
\centering
\caption{FLAME-3. Classical methods scored on 592 non-empty frames; deep methods on
the 94-frame test split. ``oracle'' = contour initialized from the GT mask.}
\begin{tabular}{llcc}
\toprule
Method & Init / family & IoU & Dice \\
\midrule
GAC                       & oracle (from GT)        & \textbf{0.365} & 0.499 \\
Kass Snakes               & oracle (from GT)        & 0.198 & 0.305 \\
GAC (U-Net seed)          & learned                 & 0.150 & 0.211 \\
U-Net                     & learned (RGB)           & 0.147 & 0.207 \\
DALS                      & learned (RGB)           & 0.122 & 0.182 \\
Deep Snake (simple)       & learned (RGB)           & 0.032 & 0.032 \\
\textbf{Deep Snake (paper)} & CenterNet $\to$ snake & 0.032 & 0.032 \\
Color floor (R$-$G)       & none                    & 0.004 & 0.007 \\
\bottomrule
\end{tabular}
\end{table}

On FLAME-3 the paper pipeline is \emph{indistinguishable from the floor} (0.032,
the empty-vs-empty score): its CenterNet detector trained to \textbf{mAP 0.000} and
detected nothing, so the snake never received a box. The whole table sits low ---
oracle-init GAC (0.365) leads only because it is handed the GT location; every
prior-free method, deep or classical, struggles because thermal-defined fire is not
recoverable from RGB (in-fire pixels measure $R{-}G\approx2.5$ vs.\ $1.9$
background; color floor $=0.004$).

\section{FLAME-1 (visible-flame)}

\begin{table}[H]
\centering
\caption{FLAME-1. Deep methods + GAC(U-Net) scored on the 301-frame test split;
color / oracle-GAC / oracle-Kass on all 2{,}001 frames (sample size in
parentheses). Deep Snake (paper) detector trained 10 epochs; snakes 30 epochs.}
\begin{tabular}{llcc}
\toprule
Method & Init / family & IoU & Dice \\
\midrule
U-Net                       & learned (RGB)           & \textbf{0.772} & 0.870 \\
DALS                        & learned (RGB)           & 0.747 & 0.854 \\
Deep Snake (simple)         & learned (RGB)           & 0.658 & 0.792 \\
GAC (U-Net seed)            & learned                 & 0.646 & 0.784 \\
GAC                         & oracle (from GT, n=2001)& 0.592 & 0.738 \\
Color floor (R$-$G, n=2001) & none                    & 0.473 & 0.625 \\
\textbf{Deep Snake (paper)} & CenterNet $\to$ snake   & 0.050 & 0.080 \\
Kass Snakes                 & oracle (from GT, n=2001)& 0.028 & 0.052 \\
\bottomrule
\end{tabular}
\end{table}

On FLAME-1 every learned method jumps dramatically --- even the \emph{color floor}
(0.473) beats every FLAME-3 method, because visible flame really is red/orange.
U-Net leads (0.772). The detector-free Deep Snake (simple) is competitive at 0.658.

\texttt{deep\_snake\_paper}, however, lands near the bottom at \textbf{0.050} --- and
this is the report's central comparison point.

\section{Why the paper pipeline is last (on both datasets)}

The decoupled detector$\to$snake design is gated by detection, and that gate
dominates the score:

\begin{itemize}
\item \textbf{FLAME-3:} detector mAP $0.000$ $\Rightarrow$ no boxes $\Rightarrow$ floor
  (0.032). Completely detection-limited.
\item \textbf{FLAME-1:} the detector now \emph{works} (mAP $0.120$, mAP@50
  $\mathbf{0.430}$ after the learning-rate fix), but at 10 epochs it fires on only
  \textbf{24\% of test frames}. On the other 76\% the snake gets no box and outputs
  empty, dragging the mean to 0.050.
\item \textbf{The snake itself is fine.} Conditioned on the detector firing,
  \texttt{deep\_snake\_paper} reaches IoU \textbf{0.206} --- 4$\times$ its
  unconditional score --- and its oracle-box validation IoU was $0.180$. The
  refinement head works; recall is the bottleneck.
\end{itemize}

By contrast, Deep Snake (simple) seeds from a \emph{dense} U-Net coarse mask, which
fires on 100\% of frames, so it never starves --- explaining the 0.658 vs.\ 0.050
gap between two methods that share the same snake refinement core.

\section{Takeaways}

\begin{enumerate}
\item \textbf{Dataset dominates.} Visible-flame FLAME-1 lifts every learned method
  far above its FLAME-3 score; thermal-defined FLAME-3 is near-unsolvable from RGB
  without an oracle.
\item \textbf{Paper-faithful $\neq$ better here.} Faithfully decoupling detection
  from refinement \emph{added} a failure mode --- detector recall --- that the
  detector-free variant avoids. Deep Snake (paper) is last on both datasets despite
  a sound snake head.
\item \textbf{The gap is recall, and it is closable.} The detector was capped at 10
  epochs (runtime) and was still improving; FLAME-1's fire boxes are also fragmented
  (median $\sim$2 dominant components/frame). More detector training and
  less-fragmented box targets are the obvious next levers --- the snake's
  given-a-box performance (0.206) sets the headroom.
\end{enumerate}

\vspace{0.5em}
\footnotesize
Sources: FLAME-3 + FLAME-1 baselines from \texttt{results/boundary\_summary.csv} and
\texttt{RESULTS.md}; Deep Snake (paper/simple) from
\texttt{results/rerun\_flame1\_20260601\_193349/} (FLAME-1) and the FLAME-3 rerun
(\texttt{deep\_snake\_paper} 0.032, detector mAP 0). \emph{Caveats:} (i) within
FLAME-1, classical/oracle rows use all 2{,}001 frames while deep rows use the 301
test frames; (ii) oracle methods are handed the GT location and are not
detection-comparable; (iii) the paper detector ran 10 epochs, below convergence.
\end{document}
